---
title: "Factor Analysis"
author: "pg611@scarletmail.rutgers.edu"
date: "2024-03-18"
output: html_document
---

```{r}
# Factor Analysis

library(psych)

layoffdataset <- read.csv("/Users/parul/OneDrive/Desktop/MVA/Layoff_Data.csv", row.names=1)

```



```{r}
attach(layoffdataset)
cov(layoffdataset[,-1])
cor(layoffdataset[,-1])  #see the highest correlating columns, like in these its agr and sps
layoffdataset[1]
```



```{r}
fit.pc <- principal(layoffdataset[-1], nfactors=4, rotate="varimax") #giving random number of factors to determine the factors
fit.pc
```


```{r}
#fit.pc <- principal(layoffdataset[-1], nfactors=3, rotate="varimax") #giving random number of factors to determine the factors
#fit.pc
```



```{r}
round(fit.pc$values, 3)
fit.pc$loadings
```


```{r}
# Loadings with more digits
for (i in c(1,3,2,4)) { print(fit.pc$loadings[[1,i]])}
```



```{r}
# Communalities
fit.pc$communality
```



```{r}
# Rotated factor scores, Notice the columns ordering: RC1, RC3, RC2 and RC4
fit.pc$scores
```



```{r}
# Play with FA utilities

fa.parallel(layoffdataset[-1]) # See factor recommendation

```




#QUESTION 1 : Decide how many Factors are ideal for your dataset



#ANSWER 1 : The ideal number of factors for a dataset is determined using techniques like parallel analysis. In this case, parallel analysis suggests that the dataset is best represented by 2 factors.





# QUESTION 2 : Explain the output for your factor model



#ANSWER 2 : The output of factor model includes :


# •	Standardized loadings: Correlations between original variables and extracted factors (RC1, RC2, RC3, RC4).


# •	SS loadings: Sum of squared loadings for each factor, indicating variance explained.


# •	Proportion Var: Proportion of total variance explained by each factor.


# •	Cumulative Var: Cumulative proportion of total variance explained.


# •	Test of Hypothesis: Assesses if the selected factors sufficiently explain the data.


# •	Fit Based on Off Diagonal Values: Indicates model fit based on off-diagonal values.





#QUESTION 3 : Show the columns that go into each factor



# ANSWER 3 : Columns Contribution to Factors:



# •	RC1: "Before_Layoffs" and "After_layoffs" contribute significantly to RC1, indicating a strong association between these variables.


# •	RC2: "lat" and "lng" contribute most to RC2, suggesting a relationship between location coordinates.


# •	RC3: "Money_Raised_in_._mil" predominantly influences RC3, indicating its strong impact on this factor.


# •	RC4: "Percentage" variable contributes primarily to RC4, suggesting it's a significant determinant of this factor.




# QUESTION 4 : Perform some visualizations using the factors



# ANSWER : Here are the visualizations below :



```{r}
fa.plot(fit.pc) # See Correlations within Factors

```


```{r}
fa.diagram(fit.pc) # Visualize the relationship

```



```{r}
vss(layoffdataset[-1]) # See Factor recommendations for a simple structure
```



```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```

