---
title: "Multiple Regression_Layoff"
author: "pg611@scarletmail.rutgers.edu"
date: "2024-04-15"
output: html_document
---

```{r }
lay_data <- read.csv("/Users/parul/OneDrive/Desktop/MVA/Layoff_Data.csv", row.names=1)

str(lay_data)

colnames(lay_data) <- c("Laidoff", "Percentage", "BeforeLayoff", "AfterLayoff", "MoneyRaised", "lat", "lng", "laidOff", "layoff")

```


```{r }
# MODEL DEVELOPMENT 


# Performing multiple regression on mtcars dataset
fit <- lm(layoff~Percentage + BeforeLayoff + AfterLayoff + MoneyRaised + lat + lng, data=lay_data) # LM- ASSUMES IT TO BE IT LINEAR, y(mpg)- many variables-output-residuals means error -median of the errors- median means real minus predicted

#show the results
```
INSIGHTS : The lm() function fits a linear regression model to the data, where the response variable (layoff) is modeled as a linear combination of the predictor variables (Percentage, BeforeLayoff, AfterLayoff, MoneyRaised, lat, lng). 



```{r }
# MODEL ACCEPTANCE

summary(fit)
#Summary has three sections. Section1: How well does the model fit the data (before Coefficients). Section2: Is the hypothesis supported? (until sifnif codes). Section3: How well does data fit the model (again).

```
INSIGHTS : The model's adjusted R-squared value is 0.4372, indicating that approximately 43.72% of the variability in layoff can be explained by the predictor variables included in the model.
The F-statistic tests the overall significance of the model and suggests that the model as a whole is statistically significant (p-value: 1.91e-05).


```{r }
#for values pr|>t|) greater than 0.05, its not considered right
fit <- lm(layoff~Percentage + BeforeLayoff + AfterLayoff + MoneyRaised + lat + lng, data=lay_data)#removed the values that were 0.05
# Useful Helper Functions
coefficients(fit)
library(GGally)
ggpairs(data=lay_data, title="Layoff Data")
confint(fit,level=0.95)#confidence interval
fitted(fit)
```


```{r }
# Residual ANALYSIS


residuals(fit)
#Anova Table
anova(fit)
vcov(fit)
cov2cor(vcov(fit))
temp <- influence.measures(fit)#get to know abt outliers of data - makes * at the end
temp
plot(fit) #current one shows outlier as it is first
```
INSIGHTS : The overall regression model is significant as indicated by a low p-value (p < 0.05).
BeforeLayoff and AfterLayoff variables are significant predictors of layoff, as their p-values are less than 0.05.
Percentage, MoneyRaised, lat, and lng variables do not appear to be significant predictors, as their p-values are greater than 0.05.



```{r }
# Assessing Outliers
library(car)
outlierTest(fit)#current one shows outlier 
leveragePlots(fit) # leverage plots
# Influential Observations
# added variable plots
avPlots(fit)
```


```{r }
# Cook's D plot
# identify D values > 4/(n-k-1)
cutoff <- 4/((nrow(mtcars)-length(fit$coefficients)-2))
plot(fit, which=4, cook.levels=cutoff)
```



```{r }
# Influence Plot
influencePlot(fit, id.method="identify", main="Influence Plot", sub="Circle size is proportial to Cook's Distance" )
# Normality of Residuals
# qq plot for studentized resid
qqPlot(fit, main="QQ Plot")
```



```{r }
# distribution of studentized residuals
library(MASS)
sresid <- studres(fit)
hist(sresid, freq=FALSE, main="Distribution of Studentized Residuals")
xfit<-seq(min(sresid),max(sresid),length=40)
yfit<-dnorm(xfit)
lines(xfit, yfit)
```



```{r }
#Non-constant Error Variance
# Evaluate homoscedasticity
# non-constant error variance test
ncvTest(fit)
# plot studentized residuals vs. fitted values
spreadLevelPlot(fit)
```


```{r }
#Multi-collinearity
# Evaluate Collinearity
vif(fit) # variance inflation factors
sqrt(vif(fit)) > 2 # problem?
```



```{r }
#Nonlinearity
# component + residual plot
crPlots(fit)
# Ceres plots
ceresPlots(fit)
#Non-independence of Errors
# Test for Autocorrelated Errors
durbinWatsonTest(fit)
```



```{r }
# Global test of model assumptions
library(gvlma)
gvmodel <- gvlma(fit)
summary(gvmodel)
fit
summary(fit)
fit1 <- fit
fit2 <- lm(layoff~Percentage + BeforeLayoff + AfterLayoff + MoneyRaised + lat + lng, data = lay_data)
# compare models
```



```{r }
# ACCURACY 

anova(fit1, fit2)
step <- stepAIC(fit, direction="both")
step$anova # display results
library(leaps)
leaps<-regsubsets(mpg~disp+hp+drat+wt+qsec,data=mtcars,nbest=10)
# view results
plot(leaps)
plot(leaps,scale="r2")
plot(leaps,scale="bic")
summary(leaps)
library(relaimpo)
calc.relimp(fit,type=c("lmg","last","first","pratt"),
            rela=TRUE)
```
INSIGHTS : The final selected model (layoff ~ BeforeLayoff + AfterLayoff) has the lowest AIC value (-102.065) among all evaluated models.
This model includes only the BeforeLayoff and AfterLayoff predictors, indicating that these variables alone provide the best balance of model complexity and explanatory power for predicting layoff.




```{r }
# Bootstrap Measures of Relative Importance (1000 samples)
boot <- boot.relimp(fit, b = 1000, type = c("lmg",
                                            "last", "first", "pratt"), rank = TRUE,
                    diff = TRUE, rela = TRUE)
booteval.relimp(boot) # print result
plot(booteval.relimp(boot,sort=TRUE)) # plot result

```
INSIGHTS : Total response variance: 0.235102
Analysis based on 50 observations.
Proportion of variance explained by the model: 50.61%.


```{r }
# ACCURACY 

summary(fit)

# PREDICTION

predict.lm(fit, data.frame(Percentage =0.5 ,BeforeLayoff=10,AfterLayoff = 8,MoneyRaised = 2000,lat = 40,lng = -75,hp=130) )

```
INSIGHTS : The function predicts the value of the response variable for a specific set of predictor variable values.
In this case, the predicted value of the response variable (layoff) is approximately 0.0895.


```{r }

```



```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```


```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```



```{r }

```

