---
title: "Principle Component Analysis"
author: "pg611@scarletmail.rutgers.edu"
date: "2024-03-04"
output: html_document
---

```{r}
library(readr)
library(lattice)
library(ggplot2)
library(ggridges)
library(ggvis)
library(ggthemes)
library(cowplot)
library(gapminder)
library(gganimate)
library(dplyr)
library(tidyverse)
library(grid)
library(gridExtra)
library(RColorBrewer)
layoff <- read_csv("/Users/parul/OneDrive/LAYOFF1.csv")
```




```{r}
str(layoff[,7:10])
```

```{r}
attach(layoff[,7:10])
```

```{r}
#Get the Correlations between the measurements
cor(layoff[,7:10][-1])
```

```{r}
# Using prcomp to compute the principal components (eigenvalues and eigenvectors). With scale=TRUE, variable means are set to zero, and variances set to one
layoff_pca <- prcomp(layoff[,7:10][,-1],scale=TRUE)
layoff_pca
summary(layoff_pca)

```

```{r}
# Eigenvalues are sdev^2
(eigen_layoff <- layoff_pca$sdev^2)
names(eigen_layoff) <- paste("PC",1:3,sep="")
eigen_layoff
sumlambdas <- sum(eigen_layoff)
sumlambdas
propvar <- eigen_layoff/sumlambdas
propvar
cumvar_layoff <- cumsum(propvar)
cumvar_layoff
matlambdas <- rbind(eigen_layoff,propvar,cumvar_layoff)
rownames(matlambdas) <- c("Eigenvalues","Prop. variance","Cum. prop. variance")
round(matlambdas,4)
summary(layoff_pca)
layoff_pca$rotation
print(layoff_pca)
```


QUESTION : Decide how many Principal Components (PCs) you want to keep and why 


ANSWER : The summary of PCA indicates the standard deviation and proportion of variance explained by each PC. In this case, the first PC (PC1) explains 67.45% of the variance, while the second PC (PC2) explains 32.53% of the variance. The cumulative proportion of variance explained by PC1 and PC2 is approximately 99.99%, indicating that these two PCs capture almost all of the variance in the data.

Based on these values, you could consider retaining two principal components (PCs) because:

PC1 and PC2 together explain almost all of the variance in the data (99.99% cumulative proportion of variance explained).
Retaining only two PCs simplifies the analysis while still capturing the majority of the variability in the original data.


```{r}
layoff_pca$x
```

QUESTION : Explain the variate representation each PCs




ANSWER 2 :

PC1: The first principal component is primarily characterized by a positive loading for the variable "Percentage" (0.1514), indicating a positive correlation between this variable and PC1. The other two variables, "Company_Size_before_Layoffs" and "Company_Size_after_layoffs," have negative loadings, indicating a negative correlation with PC1.

PC2: The second principal component has a strong negative loading for the variable "Percentage" (-0.9885), indicating a strong negative correlation between this variable and PC2. The other variables have relatively smaller loadings, with both "Company_Size_before_Layoffs" and "Company_Size_after_layoffs" having negative loadings.

PC3: The third principal component has very small loadings for all variables, suggesting that it explains very little variance in the original data compared to PC1 and PC2. The variable "Percentage" has a slightly positive loading, while the other variables have negligible loadings.

In summary, PC1 captures the overall variation in the data related to the "Percentage" variable, while PC2 captures the variation mostly in "Company_Size_before_Layoffs" and "Company_Size_after_layoffs," with an emphasis on the negative correlation with "Percentage." PC3 contributes very little to the overall variance explained by the principal components and may not be meaningful for interpretation based on the provided rotation matrix.
```{r}
layofftyp_pca <- cbind(data.frame(layoff$Laid_Off_Yes_No),layoff_pca$x)
layofftyp_pca
```

```{r}
# Means of scores 
tabmeansPC <- aggregate(layofftyp_pca[,2:4],by=list(Laid_Off_Yes_No=layoff$Laid_Off_Yes_No),mean)
tabmeansPC
tabmeansPC <- tabmeansPC[rev(order(tabmeansPC$Laid_Off_Yes_No)),]
tabmeansPC
tabfmeans <- t(tabmeansPC[,-1])
tabfmeans
colnames(tabfmeans) <- t(as.vector(tabmeansPC[1]$Laid_Off_Yes_No))
tabfmeans


```

```{r}
# Standard deviations of scores 
tabsdsPC <- aggregate(layofftyp_pca[,2:4],by=list(Laid_Off_Yes_No=layoff$Laid_Off_Yes_No),sd)
tabfsds <- t(tabsdsPC[,-1])
colnames(tabfsds) <- t(as.vector(tabsdsPC[1]$Laid_Off_Yes_No))
tabfsds
t.test(PC1~layoff$Laid_Off_Yes_No,data=layofftyp_pca)
t.test(PC2~layoff$Laid_Off_Yes_No,data=layofftyp_pca)
t.test(PC3~layoff$Laid_Off_Yes_No,data=layofftyp_pca)
```


```{r}
## F ratio tests
var.test(PC1~layoff$Laid_Off_Yes_No,data=layofftyp_pca)
var.test(PC2~layoff$Laid_Off_Yes_No,data=layofftyp_pca)
var.test(PC3~layoff$Laid_Off_Yes_No,data=layofftyp_pca)

```

```{r}
# Levene's tests (one-sided)
library(car)
(LTPC1 <- leveneTest(PC1~layoff$Laid_Off_Yes_No,data=layofftyp_pca))
(p_PC1_1sided <- LTPC1[[3]][1]/2)
(LTPC2 <- leveneTest(PC2~layoff$Laid_Off_Yes_No,data=layofftyp_pca))
(p_PC2_1sided=LTPC2[[3]][1]/2)
(LTPC3 <- leveneTest(PC3~layoff$Laid_Off_Yes_No,data=layofftyp_pca))
(p_PC3_1sided <- LTPC3[[3]][1]/2)
```

```{r}
# Plotting the scores for the first and second components
plot(layofftyp_pca$PC1, layofftyp_pca$PC2,pch=ifelse(layofftyp_pca$Laid_Off_Yes_No == "Yes",1,16),xlab="PC1", ylab="PC2", main="Layoff against values for PC1 & PC2")
abline(h=0)
abline(v=0)
legend("bottomleft", legend=c("Layoff","No Layoff"), pch=c(1,16))
plot(eigen_layoff, xlab = "Component number", ylab = "Component variance", type = "l", main = "Scree diagram")
plot(log(eigen_layoff), xlab = "Component number",ylab = "log(Component variance)", type="l",main = "Log(eigenvalue) diagram")
print(summary(layoff_pca))
diag(cov(layoff_pca$x))
xlim <- range(layoff_pca$x[,1])
layoff_pca$x[,1]
layoff_pca$x
plot(layoff_pca$x,xlim=xlim,ylim=xlim)
layoff_pca$rotation[,1]
layoff_pca$rotation
plot(layoff[,-1])
layoff_pca$x
plot(layoff_pca)
```




In addition to the correlation matrix and the summary of principal components (PCs), the scree plot also supports the decision to retain two principal components (PCs) in the analysis.



```{r}
#get the original value of the data based on PCA
center <- layoff_pca$center
scale <- layoff_pca$scale
new_layoff <- as.matrix(layoff[,7:10][,-1])
new_layoff
```


```{r}
drop(scale(new_layoff,center=center, scale=scale)%*%layoff_pca$rotation[,1])
predict(layoff_pca)[,1]
```


```{r}
layoff$Laid_Off_Yes_No <- as.factor(layoff$Laid_Off_Yes_No)
out <- sapply(1:3, function(i){plot(layoff$Laid_Off_Yes_No,layoff_pca$x[,i],xlab=paste("PC",i,sep=""),ylab="layoff")})
pairs(layoff_pca$x[,1:3], ylim = c(-6,4),xlim = c(-6,4),panel=function(x,y,...){text(x,y,layoff$Laid_Off_Yes_No)})

```


```{r}
library(factoextra)
library(FactoMineR)
library(psych)
library(corrplot)
library(devtools)
```

VISUALIZATIONS:

```{r}
# Correlation
pairs.panels(layoff[,-1],
             gap = 0,
             bg = c("red", "blue")[layoff$Laid_Off_Yes_No],
             pch=21)
```



```{r}
pairs.panels(layoff_pca$x,
             gap=0,
             bg = c("red", "blue")[layoff$Laid_Off_Yes_No],
             pch=21)

```



```{r}
fviz_eig(layoff_pca, addlabels = TRUE)
```


```{r}
fviz_pca_var(layoff_pca,col.var = "cos2",
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"),
             repel = TRUE)
```



```{r}
fviz_pca_ind(layoff_pca, col.ind = "cos2", 
             gradient.cols = c("#FFCC00", "#CC9933", "#660033", "#330033"), 
             repel = TRUE)
```



```{r}
biplot(layoff_pca)
```





```{r}
# Different PCA Method. 
res.pca <- PCA(layoff[,7:10][,-1], graph = FALSE)
print(res.pca)
```


```{r}
eig.val <- get_eigenvalue(res.pca)
eig.val

fviz_eig(res.pca, addlabels = TRUE, ylim = c(0, 50))
```


```{r}
var <- get_pca_var(res.pca)
var
```


```{r}
# Coordinates
head(var$coord)
# Cos2: quality on the factore map
head(var$cos2)
# Contributions to the principal components
head(var$contrib)
```


```{r}
# Correlation circle
fviz_pca_var(res.pca, col.var = "black")
```


```{r}
corrplot(var$cos2, is.corr=FALSE)
```


```{r}
fviz_cos2(res.pca, choice = "var", axes = 1:2)
```



```{r}
fviz_pca_var(res.pca, col.var = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"), 
             repel = TRUE # Avoid text overlapping
)
```

```{r}
fviz_pca_var(res.pca, alpha.var = "cos2")
corrplot(var$contrib, is.corr=FALSE)
# Contributions of variables to PC1
fviz_contrib(res.pca, choice = "var", axes = 1, top = 10)
# Contributions of variables to PC2
fviz_contrib(res.pca, choice = "var", axes = 2, top = 10)
fviz_pca_var(res.pca, col.var = "contrib",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07")
)
fviz_pca_var(res.pca, alpha.var = "contrib")
```

```{r}
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = layoff$Laid_Off_Yes_No, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups"
)
```

```{r}

```

```{r}
ind <- get_pca_ind(res.pca)
ind
```

```{r}
# Coordinates of individuals
head(ind$coord)
# Quality of individuals
head(ind$cos2)
# Contributions of individuals
head(ind$contrib)
```

```{r}
fviz_pca_ind(res.pca)

fviz_pca_ind(res.pca, col.ind = "cos2", 
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
)
fviz_pca_ind(res.pca, pointsize = "cos2", 
             pointshape = 21, fill = "#E7B800",
             repel = TRUE # Avoid text overlapping (slow if many points)
)
```

```{r}
fviz_pca_ind(res.pca, col.ind = "cos2", pointsize = "cos2",
             gradient.cols = c("#00AFBB", "#E7B800", "#FC4E07"),
             repel = TRUE # Avoid text overlapping (slow if many points)
)
fviz_cos2(res.pca, choice = "ind")
# Total contribution on PC1 and PC2
fviz_contrib(res.pca, choice = "ind", axes = 1:2)
```

```{r}

```

```{r}
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (nbut not "text")
             col.ind = layoff$Laid_Off_Yes_No, # color by groups
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, # Concentration ellipses
             legend.title = "Groups"
)
```

```{r}
fviz_pca_ind(res.pca, geom.ind = "point", col.ind = layoff$Laid_Off_Yes_No, 
             palette = c("#00AFBB", "#E7B800", "#FC4E07"),
             addEllipses = TRUE, ellipse.type = "confidence",
             legend.title = "Groups"
)
```

```{r}
fviz_pca_ind(res.pca,
             label = "none", # hide individual labels
             habillage = layoff$Laid_Off_Yes_No, # color by groups
             addEllipses = TRUE, # Concentration ellipses
             palette = "jco"
)
```

```{r}
fviz_pca_var(res.pca, geom.var = c("point", "text"))
# Show individuals text labels only
fviz_pca_ind(res.pca, geom.ind =  "text")
# Change the size of arrows an labels
fviz_pca_var(res.pca, arrowsize = 1, labelsize = 5, 
             repel = TRUE)
```

```{r}
fviz_pca_ind(res.pca, 
             pointsize = 3, pointshape = 21, fill = "lightblue",
             labelsize = 5, repel = TRUE)

```

```{r}
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (but not "text")
             group.ind = layoff$Laid_Off_Yes_No, # color by groups
             legend.title = "Groups",
             mean.point = FALSE)
fviz_pca_ind(res.pca,
             geom.ind = "point", # show points only (but not "text")
             group.ind = layoff$Laid_Off_Yes_No, # color by groups
             legend.title = "Groups",
             mean.point = TRUE)
fviz_pca_var(res.pca, axes.linetype = "blank")
```

```{r}
ind.p <- fviz_pca_ind(res.pca, geom = "point", col.ind = layoff$Laid_Off_Yes_No)
ggpubr::ggpar(ind.p,
              title = "Principal Component Analysis",
              subtitle = "layoff data set",
              caption = "Source: factoextra",
              xlab = "PC1", ylab = "PC2",
              legend.title = "Laid Off", legend.position = "top",
              ggtheme = theme_gray(), palette = "jco"
)

```

```{r}

```

```{r}
fviz_pca_biplot(res.pca, 
                # Individuals
                geom.ind = "point",
                fill.ind = layoff$Laid_Off_Yes_No, col.ind = "black",
                pointshape = 21, pointsize = 2,
                palette = "jco",
                addEllipses = TRUE,
                # Variables
                alpha.var ="contrib", col.var = "contrib",
                gradient.cols = "RdYlBu",
                
                legend.title = list(fill = "Laid Off", color = "Contrib",
                                    alpha = "Contrib")
)


```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

```{r}

```

